Nested Stochastic Block Models Applied to the Analysis of Single Cell Data

### 2021-07-18 manuscript.tex
added last figure

\textcolor{red}{As the size and number of communities is strictly dependent on the \emph{k}NN graph generation, we also investigated how different parameters (\emph{i.e.} number of principal components and number of neighbors) affect the partition structure (Fig SUPPBAZBAZ), again using the Crypt experiment as representative sample. We found, as a general pattern, that increasing the number of neighbors results in more granular structure at level 0, whereas increasing the number of PCs works in opposite direction. Looking at the agreement among different partitions, we found that it is consistent among different solutions, both at level 0 and level 1, with the exception of models for graph derived with low number of PC (Fig. ABRACADABRA)}
%\subsection*{PPBM?}

@@

The major drawback of adopting this strategy is the substantial increase of runtimes. As observed, model minimisation may be hundred times slower than the extremely fast Leiden approach. It should be noted that \texttt{schist} initialises multiple models that are treated by multiple concurrent processes. \texttt{graph-tool} itself supports CPU-level parallelisation for some of its tasks. These optimisations are well suited for clustered computing infrastructure. Further development, possibly including GPU-level parallelisation, is surely required to accomodate the large size of datasets that are being produced.

@@

>>>>

\textcolor{red}{As the size and number of communities is strictly dependent on the \emph{k}NN graph generation, we investigated how different parameters (\emph{i.e.} number of principal components and number of neighbors) affect the partition structure (Fig SUPPBAZBAZ). We found, as a general pattern, that increasing the number of neighbors results in more granular structure at level 0, whereas increasing the number of PCs works in opposite direction. Looking at the agreement among different partitions, we found that it is consistent among different solutions, both at level 0 and level 1, with the exception of models for graph derived with low number of PC (Fig. ABRACADABRA)}
%\subsection*{PPBM?}

@@

The major drawback of adopting this strategy is the substantial increase of runtimes. As observed, model minimisation is many times slower than the extremely fast Leiden approach. It should be noted that \texttt{schist} initialises multiple models that are treated by multiple concurrent processes. \texttt{graph-tool} itself supports CPU-level parallelisation for some of its tasks. These optimisations are well suited for clustered computing infrastructure. Further development, possibly including GPU-level parallelisation, is surely required to accomodate the large size of datasets that are being produced.

@@

\subsection*{Analysis of crypt cells data}
\textcolor{red}{Count matrix for untreated crypt cells (GSM3308718) was downloaded from GEO. Cells with less than 200 genes and genes detected in less than 2 cells were excluded from the analysis. After normalization and log-transformation, highly variable genes were selected with a cutoff on the mean expression equal to 0.05. Rare subpopulations were first highlighted using the gene signatures published in \cite{Fa_Wei_Zhou_Johnston_2021}:
}
\begin{itemize}
\item R1\_1: Cd8a, Cd3g, Ccl5, Gzma, Gzmb, RGs1, Nkg7, Cd7, Fcer1g
\item R1\_2: H2-Aa, H2-Ab1, H2-Eb1, Cd74, Ly6d, Ebf1, Cd79a, Mef2c
\item R2: Krt18, Cd24a, Adh1, Cystm1, Aldh2, Dclk1, Sh2d6, Rgs13, Hck, Trpm5
\item R3: Alas2, Hbb-bs, Hba-a1, Hbb-bt
\end{itemize}

### 2021-07-18 supplementary.tex
last complex figure
added figure for tuft cells
added last figure

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{hist_label_second.png}
\caption[]{Distribution of second choices in label transfer. For each dataset in which we assessed accuracy of label transfer by SMB, we collect the rank of affinities of the correct class when an assignment could not be performed (\emph{i.e.} the highest affinity was for the "Unknonwn" label)  }\label{hist_label_second}
\end{figure}
\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.9\textheight]{FigureOptimalLevel.png}
\caption[]{Choice of the optimal hierarchy level. Datasets analysed for choice of the most relevant hierarchy level are represented here, from top to bottom in the same order as in Table \ref{table_optimal}. For each dataset we report the UMAP embedding coloured by the annotation given in the corresponding manuscript, the profile of modularity $Q$ at different level of the hierarchy and the UMAP embeddings coloured by the level according to $i^k$ or $i^Q$. The dashed line in modularity plots represents the modularity calculated using the annotation from the corresponding manuscripts.}\label{FigureOptimalLevel}
\end{figure}

>>>>

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{hist_label_second.png}
\caption[]{Distribution of second choices in label transfer. For each dataset in which we assessed accuracy of label transfer by SBM, we collect the rank of affinities of the correct class when an assignment could not be performed (\emph{i.e.} the highest affinity was for the "Unknonwn" label)  }\label{hist_label_second}
\end{figure}
\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=1\textwidth,height=\textheight]{FigureOptimalLevel.pdf}
\vspace*{-80pt}
\caption[]{Choice of the optimal hierarchy level. Datasets analysed for choice of the most relevant hierarchy level are represented here, from top to bottom in the same order as in Table \ref{table_optimal}. For each dataset we report the UMAP embedding coloured by the annotation given in the corresponding manuscript, the profile of modularity $Q$ at different level of the hierarchy and the UMAP embeddings coloured by the level according to $i^k$ or $i^Q$. The dashed line in modularity plots represents the modularity calculated using the annotation from the corresponding manuscripts.}\label{FigureOptimalLevel}
\end{figure}

\clearpage
\begin{figure}[H]
\centering
\includegraphics[keepaspectratio,width=1\textwidth,height=\textheight]{Figure_Rare_Cells_Supp.pdf}
%\vspace*{-40pt}
\caption[]{Identification of rare populations. FOOBAR} \label{Figure_Rare_Supp}
\end{figure}

\clearpage
\begin{figure}[H]
\centering
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel0.pdf} 
        \caption{Chromium 10x} \label{cm_10x}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel1.pdf} 
        \caption{iCell8} \label{cm_icell8}
    \end{subfigure}
    \hfill
%    \vspace{40pt}
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel2.pdf} 
        \caption{Quartz-seq2} \label{cm_quartz}
    \end{subfigure}

    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel3.pdf} 
        \caption{MARS-seq} \label{cm_mars}
    \end{subfigure}
%    \vspace{40pt}
   \hfill
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel4.pdf} 
        \caption{Crypt cells} \label{cm_crypt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel5.pdf} 
        \caption{Planaria} \label{cm_planaria}
    \end{subfigure}
    \vspace{40pt}

%\vspace*{-40pt}
\caption[]{Consistency at Level 0} \label{Figure_Clustermap_L0}
\end{figure}
